
---
title: "BRFSS Modeling"
author: "Steven S., Ben K., and Chris R."
date: "2025-04-18"
output: html_document
---

```{r}
library(ggplot2)
library(nnet) # Load for multinomial logistic regression
library(ranger) # Fast Random Forest engine
library(yardstick)
```

##########################################################################New Code for Initial Modeling############################################################################################################

## Creating Categorical Target Variable ('ment14d_cat')- CR
```{r}
trainimputeFinal <- trainimputeFinal %>%
  mutate(ment14d_cat = case_when(
    ment14d == 1 ~ "None",         # 0 days
    ment14d == 2 ~ "Occasional",   # 1–13 days
    ment14d == 3 ~ "Frequent",     # 14+ days
    TRUE ~ NA_character_
  ))

testimputeFinal <- testimputeFinal %>%
  mutate(ment14d_cat = case_when(
    ment14d == 1 ~ "None",
    ment14d == 2 ~ "Occasional",
    ment14d == 3 ~ "Frequent",
    TRUE ~ NA_character_
  ))

trainimputeFinal$ment14d_cat <- factor(trainimputeFinal$ment14d_cat, levels = c("None", "Occasional", "Frequent"))
testimputeFinal$ment14d_cat <- factor(testimputeFinal$ment14d_cat, levels = c("None", "Occasional", "Frequent"))
```


```{r}
train_clean <- trainimputeFinal %>%
  mutate(across(where(is.integer), as.numeric)) %>%
  select(-ment14d)  # Remove the original numeric target to prevent leakage

test_clean <- testimputeFinal %>%
  mutate(across(where(is.integer), as.numeric)) %>%
  select(-ment14d)
```


# Tidymodels Recipe for Scaling/Centering - SS
```{r}
recipe <- recipe(ment14d_cat ~ ., data = train_clean) %>%  # Use cleaned training data
  update_role(menthlth, new_role = "id") %>%  # Excludes menth1th
  step_impute_mean(all_numeric_predictors()) %>% # Impute missing nummeric value
  step_upsample(ment14d_cat) %>%
  step_center(all_numeric_predictors()) %>%
  step_scale(all_numeric_predictors())

preparedRecipe <- prep(recipe, training = train_clean)

traindataFinal <- bake(preparedRecipe, new_data = NULL)
testdataFinal <- bake(preparedRecipe, new_data = test_clean)

dim(traindataFinal)
dim(testdataFinal)
head(traindataFinal)
```

```{r}
table(traindataFinal$ment14d_cat)
```

```{r}
sum(is.na(traindataFinal$primins1_88))
```


## Principle Component Analysis (PCA) on Scaled Data - CR
```{r}
pca_input_data <- traindataFinal %>%
  filter(if_all(everything(), ~ !is.na(.) & is.finite(.))) # Filter out rows with any NA or infinite values

pca_input <- pca_input_data %>% 
  select(where(is.numeric)) %>% # Keep only numeric columns for PCA
  scale() # Standardize numeric features

pca_result <- prcomp(pca_input, center = TRUE, scale. = TRUE) #PCA with centering and scaling

plot(pca_result, type = "l", main = "Scree Plot") #Visualize variance by each principle component
summary(pca_result) #Output cumulative variance to decide how many components to retain
```

```{r}
pca_scores <- as.data.frame(pca_result$x) # Get PCA scores
pca_scores$ment14d_cat <- pca_input_data$ment14d_cat # Add original outcome category to PCA scores
```

```{r}

ggplot(pca_scores, aes(x = PC1, y = PC2, color = ment14d_cat)) +
  geom_point(alpha = 0.5) +
  labs(title = "PCA: PC1 vs PC2 colored by ment14d_cat") +
  theme_minimal()
```

```{r}
set.seed(123)
kmeans_result <- kmeans(pca_scores[, 1:2], centers = 3)  # Using PC1 & PC2 # Run k-means clustering with k = 3 on PC1 & PC2

pca_scores$cluster <- as.factor(kmeans_result$cluster) # Assign labels
```

```{r}
ggplot(pca_scores, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(alpha = 0.5) +
  labs(title = "K-means Clusters (k=3) on PCA Components") +
  theme_minimal()
```

```{r}
table(Actual = pca_scores$ment14d_cat, Cluster = pca_scores$cluster) # Compare clusters to original categories


############################################### Modeling ########################################################

```{r}
multi_model <- multinom_reg(mode = "classification") %>% # Set up multinomial logistic model
  set_engine("nnet")
```

```{r}
multi_wf <- workflow() %>% # Create workflow with recipe and multinomial model
  add_model(multi_model) %>%
  add_recipe(recipe)
```

```{r}
multi_fit <- fit(multi_wf, data = traindataFinal) # Fit model using preprocessed training data
```

```{r}
multi_preds <- predict(multi_fit, new_data = testdataFinal) # Predict class labels for test set
```

```{r}
multi_probs <- predict(multi_fit, new_data = testdataFinal, type = "prob") # Predict class probabilities
```

```{r}
test_results <- bind_cols(testdataFinal, multi_preds) %>% # Combine predictions with test set
  rename(pred_class = .pred_class) # Rename predicted class
```

```{r}
accuracy(test_results, truth = ment14d_cat, estimate = pred_class) # Calculate classification accuracy on test set
```

```{r}
conf_mat(test_results, truth = ment14d_cat, estimate = pred_class) # Confusion matrix to evaluate prediction breakdown across categories
```
```{r}
set.seed(123)
folds <- vfold_cv(train_clean, v = 5, strata = ment14d_cat) # Create 5-fold cross validation with stratification on target variable

```

```{r}
multi_model <- multinom_reg(mode = "classification") %>% # Define multinomial logistic regression model for classification
  set_engine("nnet")

multi_wf <- workflow() %>% # Create workflow using formula interface with multinomial model
  add_model(multi_model) %>%
  add_formula(ment14d_cat ~ .)

```

```{r}
cv_results <- fit_resamples( # Perform 5-fold cross validation and save prediction
  multi_wf,
  resamples = folds,
  metrics = metric_set(yardstick::accuracy),
  control = control_resamples(save_pred = TRUE)
)

```

```{r}
collect_metrics(cv_results) # Mean accuracy across cross validation folds
```


```{r}
rf_model <- rand_forest( # Set up model with hyperparameters to tune
  mode = "classification",
  mtry = tune(),
  trees = 100,
  min_n = tune()
) %>%
  set_engine("ranger", importance = "impurity") # Define Random Forest classification model using ranger engine
```

```{r}
rf_wf <- workflow() %>% # Create workflow with preprocessing recipe and Random Forest model
  add_model(rf_model) %>%
  add_recipe(recipe)
```

```{r}
set.seed(123) # Set up 5-fold cross-validation with stratified sampling on outcome
folds <- vfold_cv(train_clean, v = 5, strata = ment14d_cat) 
```

```{r}
rf_results <- tune_grid( # Tune mtry and min_n using cross-validation
  rf_wf,
  resamples = folds,
  grid = 3,
  metrics = metric_set(yardstick::accuracy),
  control = control_grid(verbose = TRUE)
)
```

```{r}
saveRDS(rf_results, "rf_results.rds")
```

```{r}
file.info("rf_results.rds")
```


```{r}
best_rf <- select_best(x = rf_results, metric = "accuracy")  # Select hyperparameters with highest accuracy
```

```{r}
rf_model_final <- rand_forest(
  mode = "classification",
  mtry = best_rf$mtry,
  trees = 100,
  min_n = best_rf$min_n
) %>%
  set_engine("ranger", importance = "impurity")
```

```{r}
final_rf_wf <- workflow() %>%
  add_model(rf_model_final) %>%
  add_recipe(recipe)
```


```{r}
final_rf_fit <- fit(final_rf_wf, data = traindataFinal)
```

```{r}
rf_preds <- predict(final_rf_fit, new_data = testdataFinal) # Predict classes on test set using finalized Random Forest model

rf_probs <- predict(final_rf_fit, new_data = testdataFinal, type = "prob") # Predict probabilities
```

```{r}
rf_results_final <- bind_cols(testdataFinal, rf_preds) %>% # Combine predictions with test set
  rename(pred_class = .pred_class)

accuracy(rf_results_final, truth = ment14d_cat, estimate = pred_class) # Accuracy

conf_mat(rf_results_final, truth = ment14d_cat, estimate = pred_class) # Confusion matrix
```

```{r}
rf_results_final <- bind_cols(testdataFinal, rf_preds)

rf_results_final %>%
  metrics(truth = ment14d_cat, estimate = .pred_class)
```

```{r}
rf_results_final %>%
  metrics(truth = ment14d_cat, estimate = .pred_class)
```

```{r}
rf_results_final %>%
  group_by(ment14d_cat) %>%
  yardstick::precision(truth = ment14d_cat, estimate = .pred_class)
```

```{r}
rf_results_final %>%
  group_by(ment14d_cat) %>%
  yardstick::recall(truth = ment14d_cat, estimate = .pred_class)
```

```{r}
conf_mat(rf_results_final, truth = ment14d_cat, estimate = .pred_class) %>%
  autoplot(type = "heatmap") +
  ggtitle("Random Forest Confusion Matrix (Test Set)") +
  theme_minimal()
```

```{r}
final_rf_fit %>% # Show variable importance
  extract_fit_parsnip() %>% 
  vip::vip(num_features = 15)
```
