---
title: "EDA"
author: "Christopher Rodgers"
date: "2025-04-02"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gtsummary)
library(forcats)
library(naniar)
library(corrplot)
library(stats)
library(stringr)
library(car)

```


#import data
```{r}
# Code below imports the data to any person's machine
if (!dir.exists("C:/temp")) {
    dir.create("C:/temp")
} # If statement creates temp folder on machine

zipUrl <- "https://www.cdc.gov/brfss/annual_data/2023/files/LLCP2023XPT.zip"
zipPath <- path.expand("C:/temp/LLCP2023XPT.zip")
if (!file.exists(zipPath)) {
    download.file(zipUrl, destfile = zipPath)
} #Download the file

extractPath <- "C:/temp"
sasFile <- file.path(extractPath, "LLCP2023.XPT") 
if (!file.exists(sasFile)) {
    unzip(zipPath, exdir = extractPath)
}
list.files(extractPath) # Extract, unzip, and make sure file exists

if (file.exists(sasFile)) {
    brfss_raw <- read_xpt(sasFile)
    head(brfss_raw)
} else {
    message("File does not exist, check download or unzip process.")
} # Read file into R after specifying file variable

```

#column names
```{r}
names(brfss_raw)
```
#rename variables
```{r}
brfss_subset <- brfss_raw %>%
  select(
    ment14d   = `_MENT14D`,
    hlthpl1   = `_HLTHPL1`,
    persdoc3  = PERSDOC3,
    medcost1  = MEDCOST1,
    checkup1  = CHECKUP1,
    primins1  = PRIMINS1,
    addepev3  = ADDEPEV3,
    acedeprs  = ACEDEPRS,
    decide    = DECIDE,
    exerany2  = EXERANY2,
    pa150r4   = `_PA150R4`,
    sex       = `_SEX`,
    age80     = `_AGE80`,
    race      = `_RACE`,
    educa     = EDUCA,
    employ1   = EMPLOY1,
    marital   = MARITAL,
    incomg1   = `_INCOMG1`,
    numadult  = NUMADULT,
    state     = `_STATE`
  )

```

#view data type
```{r}
glimpse(brfss_subset)

```
#recode
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    hlthpl1   = na_if(hlthpl1, 7),
    hlthpl1   = na_if(hlthpl1, 9),
    
    persdoc3  = na_if(persdoc3, 7),
    persdoc3  = na_if(persdoc3, 9),
    
    medcost1  = na_if(medcost1, 7),
    medcost1  = na_if(medcost1, 9),
    
    checkup1  = na_if(checkup1, 7),
    checkup1  = na_if(checkup1, 9),
    
    primins1  = na_if(primins1, 7),
    primins1  = na_if(primins1, 9),
    
    exerany2  = na_if(exerany2, 7),
    exerany2  = na_if(exerany2, 9),
    
    pa150r4   = na_if(pa150r4, 7),
    pa150r4   = na_if(pa150r4, 9)
  )

```

```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    race      = na_if(race, 77),
    race      = na_if(race, 99),
    
    educa     = na_if(educa, 9),
    
    employ1   = na_if(employ1, 9),
    
    marital   = na_if(marital, 9),
    
    incomg1   = na_if(incomg1, 77),
    incomg1   = na_if(incomg1, 99),
    
    sex       = na_if(sex, 7),
    sex       = na_if(sex, 9)
  )

```

```{r}
brfss_subset <- brfss_subset %>% #SS 4.2 - Is this needed?
  mutate(
    ment14d = ifelse(ment14d == 88, 0, ment14d),
    ment14d = na_if(ment14d, 77),
    ment14d = na_if(ment14d, 99)
  )

```

#check ment14d
```{r}
table(brfss_subset$ment14d, useNA = "ifany")

```
#check hlthpl1, race, incomg1
```{r}
table(brfss_subset$hlthpl1, useNA = "ifany")
table(brfss_subset$race, useNA = "ifany")
table(brfss_subset$incomg1, useNA = "ifany")

```
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    ment14d = na_if(ment14d, 9)
  )

```

```{r}
brfss_subset <- brfss_subset %>% 
  mutate(
    race = as.numeric(race),
    race = na_if(race, 77), #Do any values meet this criteria?
    race = na_if(race, 99) #Do any values meet this criteria?
  )
```

```{r}
brfss_subset <- brfss_subset %>% # SS 4.2 - Duplicate?
  mutate(
    race = as.numeric(race),
    race = na_if(race, 77),
    race = na_if(race, 99)
  )

```
#recheck
```{r}
table(brfss_subset$race, useNA = "ifany")
```
#try to fix race again
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    race = na_if(race, 7),
    race = na_if(race, 8),
    race = na_if(race, 9)
  )

```

```{r}
table(brfss_subset$race, useNA = "ifany")

```
#convert yes/no binary variables to factors
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    hlthpl1   = factor(hlthpl1, levels = c(1, 2), labels = c("Yes", "No")),
    persdoc3  = factor(persdoc3, levels = c(1, 2), labels = c("Yes", "No")),
    medcost1  = factor(medcost1, levels = c(1, 2), labels = c("Yes", "No")),
    primins1  = factor(primins1, levels = c(1, 2), labels = c("Yes", "No")),
    exerany2  = factor(exerany2, levels = c(1, 2), labels = c("Yes", "No"))
  )

```

#check conversion
```{r}
table(brfss_subset$hlthpl1, useNA = "ifany") #Noting class imbalance here, this is good!

```
```{r}
table(brfss_subset$exerany2, useNA = "ifany")

```
#convert multi level categorical variables to labeled factors
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    race = factor(race,
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c("White only", "Black only", "American Indian/Alaska Native", "Asian only", "Other single race", "Multiracial")
    ),
    
    educa = factor(educa,
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c("Never attended school", "Grades 1-8", "Grades 9-11", "High school grad", "Some college", "College grad")
    ),
    
    incomg1 = factor(incomg1,
      levels = c(1, 2, 3, 4, 5, 6, 7, 8),
      labels = c("Less than $10k", "$10k–$15k", "$15k–$20k", "$20k–$25k", "$25k–$35k", "$35k–$50k", "$50k–$75k", "Over $75k")
    ),
    
    marital = factor(marital,
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c("Married", "Divorced", "Widowed", "Separated", "Never married", "Unmarried couple")
    )
  )

```

#remaining categorical variables
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    sex = factor(sex,
      levels = c(1, 2),
      labels = c("Male", "Female")
    ),
    
    employ1 = factor(employ1,
      levels = c(1, 2, 3, 4, 5, 6, 7, 8),
      labels = c("Employed for wages", "Self-employed", "Out of work >1yr", "Out of work <1yr", "Homemaker", "Student", "Retired", "Unable to work")
    ),
    
    pa150r4 = factor(pa150r4,
      levels = c(1, 2),
      labels = c("Meets guidelines", "Does not meet")
    )
  )

```

#check to ensure recoded factor variables
```{r}
table(brfss_subset$race, useNA = "ifany")
table(brfss_subset$educa, useNA = "ifany")
table(brfss_subset$sex, useNA = "ifany")
table(brfss_subset$employ1, useNA = "ifany")

```
#summary table
```{r}
brfss_subset %>% #Are we excluding any variables in the select statement. Otherwise, it may be redunant?
  select(ment14d, hlthpl1, persdoc3, medcost1, checkup1, primins1,
         addepev3, acedeprs, decide, exerany2, pa150r4, sex, age80,
         race, educa, employ1, marital, incomg1, numadult) %>%
  tbl_summary(
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    missing = "no"
  ) %>%
  bold_labels()

```

#stratified summary table hlthpl1
```{r}
brfss_subset %>%
  select(ment14d, persdoc3, medcost1, checkup1, primins1, addepev3,
         acedeprs, decide, exerany2, pa150r4, sex, age80, race, educa,
         employ1, marital, incomg1, numadult, hlthpl1) %>%
  tbl_summary(
    by = hlthpl1,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    missing = "no"
  ) %>%
  add_p() %>%
  bold_labels()

```

#stratified summary table _ment14d
```{r}
brfss_subset %>%
  select(ment14d, persdoc3, medcost1, checkup1, primins1, addepev3,
         acedeprs, decide, exerany2, pa150r4, sex, age80, race, educa,
         employ1, marital, incomg1, numadult, hlthpl1) %>%
  tbl_summary(
    by = ment14d,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    missing = "no"
  ) %>%
  add_p() %>%
  bold_labels()

```

#race distribution bar plot
```{r}
ggplot(brfss_subset, aes(x = race)) +
  geom_bar() +
  labs(title = "Race Distribution", x = "Race", y = "Count") +
  theme_minimal()

```
#ment14d histogram
```{r}
ggplot(brfss_subset, aes(x = ment14d)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  labs(title = "Distribution of Poor Mental Health Days", x = "Days", y = "Count") +
  theme_minimal()

```
#ment14d by hlthpl1 boxplot
```{r}
ggplot(brfss_subset, aes(x = hlthpl1, y = ment14d)) +
  geom_boxplot() +
  labs(title = "Poor Mental Health Days by Insurance Status", x = "Insurance", y = "Mental Health Days") +
  theme_minimal()

```
#exerany2 bar plot
```{r}
ggplot(brfss_subset, aes(x = exerany2)) +
  geom_bar() +
  labs(title = "Exercise Participation", x = "Exercised in Past 30 Days", y = "Count") +
  theme_minimal()

```
#incomg1
```{r}
ggplot(brfss_subset, aes(x = incomg1)) +
  geom_bar() +
  labs(title = "Income Group Distribution", x = "Income Group", y = "Count") +
  theme_minimal()

```
#missingness plot
```{r}
gg_miss_var(brfss_subset) +
  labs(title = "Missingness by Variable")

```
#collinearity check
```{r}
brfss_numeric <- brfss_subset %>%
  select(ment14d, age80, numadult) %>% # SS 4.2 - We may want to consider removing. Corrplots are ideally for continuous variables, and the response should probably be excluded.
  na.omit()

cor_matrix <- cor(brfss_numeric)
corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.8)

```
# Basic Regression Analysis for response
```{r}
lmModel <- lm(ment14d ~ hlthpl1, data = brfss_subset)
summary(lmModel)

```
# Basic Regression with a few more demographic variables
```{r}
lmModelAddition <- lm(ment14d ~ hlthpl1 + sex + incomg1, data = brfss_subset)
summary(lmModelAddition)
vif(lmModelAddition) # Low multicollinearity - this is good!
```
# Chi Squared Test for days since last checkup and medical cost
```{r}
tableChiSq <- table(brfss_subset$medcost1, brfss_subset$checkup1)

chiResult <- chisq.test(tableChiSq)
print(chiResult) # SS 4.2 - Indicates checkups related to medical costs. May want to examine further
```

#PCA biplot
```{r}
pca_result <- prcomp(brfss_numeric, scale. = TRUE)
biplot(pca_result, main = "PCA Biplot (Centered & Scaled Variables)")

```

