---
title: "EDA"
author: "Christopher Rodgers"
date: "2025-04-02"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(haven)
library(dplyr)
library(tidyr)
library(ggplot2)
library(gtsummary)
library(forcats)
library(naniar)
library(corrplot)
library(stats)
library(stringr)
library(car)
library(reshape2)
library(viridis)
library(GGally)
```


# Import data
```{r}
# Code below imports the data to any person's machine
if (!dir.exists("C:/temp")) {
    dir.create("C:/temp")
} # If statement creates temp folder on machine

zipUrl <- "https://www.cdc.gov/brfss/annual_data/2023/files/LLCP2023XPT.zip"
zipPath <- path.expand("C:/temp/LLCP2023XPT.zip")
if (!file.exists(zipPath)) {
    download.file(zipUrl, destfile = zipPath)
} #Download the file

extractPath <- "C:/temp"
sasFile <- file.path(extractPath, "LLCP2023.XPT") 
if (!file.exists(sasFile)) {
    unzip(zipPath, exdir = extractPath)
}
list.files(extractPath) # Extract, unzip, and make sure file exists

if (file.exists(sasFile)) {
    brfss_raw <- read_xpt(sasFile)
    head(brfss_raw)
} else {
    message("File does not exist, check download or unzip process.")
} # Read file into R after specifying file variable

```

# Column names
```{r}
names(brfss_raw)
```
# Rename variables
```{r}
brfss_subset <- brfss_raw %>%
  select(
    ment14d   = `_MENT14D`,
    hlthpl1   = `_HLTHPL1`,
    persdoc3  = PERSDOC3,
    medcost1  = MEDCOST1,
    checkup1  = CHECKUP1,
    primins1  = PRIMINS1,
    addepev3  = ADDEPEV3,
    acedeprs  = ACEDEPRS,
    decide    = DECIDE,
    exerany2  = EXERANY2,
    pa150r4   = `_PA150R4`,
    sex       = `_SEX`,
    age80     = `_AGE80`,
    race      = `_RACE`,
    educa     = EDUCA,
    employ1   = EMPLOY1,
    marital   = MARITAL,
    incomg1   = `_INCOMG1`,
    numadult  = NUMADULT,
    state     = `_STATE`
  )

```

# View data type
```{r}
glimpse(brfss_subset)
```
# Recode
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    hlthpl1   = na_if(hlthpl1, 7),
    hlthpl1   = na_if(hlthpl1, 9),
    
    persdoc3  = na_if(persdoc3, 7),
    persdoc3  = na_if(persdoc3, 9),
    
    medcost1  = na_if(medcost1, 7),
    medcost1  = na_if(medcost1, 9),
    
    checkup1  = na_if(checkup1, 7),
    checkup1  = na_if(checkup1, 9),
    
    primins1  = na_if(primins1, 7),
    primins1  = na_if(primins1, 9),
    
    exerany2  = na_if(exerany2, 7),
    exerany2  = na_if(exerany2, 9),
    
    pa150r4   = na_if(pa150r4, 7),
    pa150r4   = na_if(pa150r4, 9)
  )

```

```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    race      = na_if(race, 77),
    race      = na_if(race, 99),
    
    educa     = na_if(educa, 9),
    
    employ1   = na_if(employ1, 9),
    
    marital   = na_if(marital, 9),
    
    incomg1   = na_if(incomg1, 77),
    incomg1   = na_if(incomg1, 99),
    
    sex       = na_if(sex, 7),
    sex       = na_if(sex, 9)
  )

```

```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    menthlth = brfss_raw$MENTHLTH,
    ment14d = ifelse(menthlth == 88, 0, menthlth),
    ment14d = na_if(ment14d, 9),
    ment14d = na_if(ment14d, 77),
    ment14d = na_if(ment14d, 99),
    ment14d_cat = case_when(
      ment14d == 0 ~ "None",
      ment14d >= 1 & ment14d <= 13 ~ "Occasional",
      ment14d >= 14 ~ "Frequent",
      TRUE ~ NA_character_
    ),
    ment14d_cat = factor(ment14d_cat, levels = c("None", "Occasional", "Frequent"))
  )
```

```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    ment14d = na_if(ment14d, 9)
  )
```

# Check ment14d
```{r}
table(brfss_subset$ment14d, useNA = "ifany")

```

```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    ment14d_cat = case_when(
      ment14d == 0 ~ "None",
      ment14d >= 1 & ment14d <= 13 ~ "Occasional",
      ment14d >= 14 ~ "Frequent",
      TRUE ~ NA_character_
    ),
    ment14d_cat = factor(ment14d_cat, levels = c("None", "Occasional", "Frequent"))
  )

```

```{r}
table(brfss_subset$ment14d_cat, useNA = "ifany")
```

# Check hlthpl1, race, incomg1
```{r}
table(brfss_subset$hlthpl1, useNA = "ifany")
table(brfss_subset$race, useNA = "ifany")
table(brfss_subset$incomg1, useNA = "ifany")

```
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    ment14d = na_if(ment14d, 9)
  )

```

```{r}
brfss_subset <- brfss_subset %>% 
  mutate(
    race = as.numeric(race),
    race = na_if(race, 77), 
    race = na_if(race, 99)
  )

```

# Recheck
```{r}
table(brfss_subset$race, useNA = "ifany")
```

# Final fix for Race
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    race = na_if(race, 7),
    race = na_if(race, 8),
    race = na_if(race, 9)
  )

```

```{r}
table(brfss_subset$race, useNA = "ifany")

```
# Convert yes/no binary variables to factors
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    hlthpl1   = factor(hlthpl1, levels = c(1, 2), labels = c("Yes", "No")),
    persdoc3  = factor(persdoc3, levels = c(1, 2), labels = c("Yes", "No")),
    medcost1  = factor(medcost1, levels = c(1, 2), labels = c("Yes", "No")),
    primins1  = factor(primins1, levels = c(1, 2), labels = c("Yes", "No")),
    exerany2  = factor(exerany2, levels = c(1, 2), labels = c("Yes", "No"))
  )

```

# Check conversion
```{r}
table(brfss_subset$hlthpl1, useNA = "ifany") #Noting class imbalance

```
```{r}
table(brfss_subset$exerany2, useNA = "ifany")

```
# Convert multi level categorical variables to labeled factors
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    race = factor(race,
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c("White only", "Black only", "American Indian/Alaska Native", "Asian only", "Other single race", "Multiracial")
    ),
    
    educa = factor(educa,
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c("Never attended school", "Grades 1-8", "Grades 9-11", "High school grad", "Some college", "College grad")
    ),
    
    incomg1 = factor(incomg1,
      levels = c(1, 2, 3, 4, 5, 6, 7, 8),
      labels = c("Less than $10k", "$10k–$15k", "$15k–$20k", "$20k–$25k", "$25k–$35k", "$35k–$50k", "$50k–$75k", "Over $75k")
    ),
    
    marital = factor(marital,
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c("Married", "Divorced", "Widowed", "Separated", "Never married", "Unmarried couple")
    )
  )

```

# Remaining categorical variables
```{r}
brfss_subset <- brfss_subset %>%
  mutate(
    sex = factor(sex,
      levels = c(1, 2),
      labels = c("Male", "Female")
    ),
    
    employ1 = factor(employ1,
      levels = c(1, 2, 3, 4, 5, 6, 7, 8),
      labels = c("Employed for wages", "Self-employed", "Out of work >1yr", "Out of work <1yr", "Homemaker", "Student", "Retired", "Unable to work")
    ),
    
    pa150r4 = factor(pa150r4,
      levels = c(1, 2),
      labels = c("Meets guidelines", "Does not meet")
    )
  )

```

# Check to ensure recoded factor variables
```{r}
table(brfss_subset$race, useNA = "ifany")
table(brfss_subset$educa, useNA = "ifany")
table(brfss_subset$sex, useNA = "ifany")
table(brfss_subset$employ1, useNA = "ifany")

```
# Summary table
```{r}
brfss_subset %>% 
  select(ment14d, hlthpl1, persdoc3, medcost1, checkup1, primins1,
         addepev3, acedeprs, decide, exerany2, pa150r4, sex, age80,
         race, educa, employ1, marital, incomg1, numadult) %>%
  tbl_summary(
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    missing = "no"
  ) %>%
  bold_labels()

```

# Stratified summary table hlthpl1
```{r}
brfss_subset %>%
  select(ment14d, persdoc3, medcost1, checkup1, primins1, addepev3,
         acedeprs, decide, exerany2, pa150r4, sex, age80, race, educa,
         employ1, marital, incomg1, numadult, hlthpl1) %>%
  tbl_summary(
    by = hlthpl1,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    missing = "no"
  ) %>%
  add_p() %>%
  bold_labels()

```

# Stratified summary table _ment14d
```{r}
brfss_subset %>%
  select(ment14d, persdoc3, medcost1, checkup1, primins1, addepev3,
         acedeprs, decide, exerany2, pa150r4, sex, age80, race, educa,
         employ1, marital, incomg1, numadult, hlthpl1) %>%
  tbl_summary(
    by = ment14d,
    statistic = list(all_continuous() ~ "{mean} ({sd})",
                     all_categorical() ~ "{n} ({p}%)"),
    missing = "no"
  ) %>%
  add_p() %>%
  bold_labels()
```

# Race distribution bar plot
```{r}
ggplot(brfss_subset, aes(x = race)) +
  geom_bar() +
  labs(title = "Race Distribution", x = "Race", y = "Count") +
  theme_minimal()
```
# Ment14d histogram
```{r}
ggplot(brfss_subset, aes(x = ment14d)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  labs(title = "Distribution of Poor Mental Health Days", x = "Days", y = "Count") +
  theme_minimal()

```
# Ment14d by hlthpl1 boxplot
```{r}
ggplot(brfss_subset, aes(x = hlthpl1, y = ment14d)) +
  geom_boxplot() +
  labs(title = "Poor Mental Health Days by Insurance Status", x = "Insurance", y = "Mental Health Days") +
  theme_minimal()

```
#exerany2 bar plot
```{r}
ggplot(brfss_subset, aes(x = exerany2)) +
  geom_bar() +
  labs(title = "Exercise Participation", x = "Exercised in Past 30 Days", y = "Count") +
  theme_minimal()

```
#incomg1
```{r}
ggplot(brfss_subset, aes(x = incomg1)) +
  geom_bar() +
  labs(title = "Income Group Distribution", x = "Income Group", y = "Count") +
  theme_minimal()

```
#missingness plot
```{r}
gg_miss_var(brfss_subset) +
  labs(title = "Missingness by Variable")

```
#collinearity check
```{r}
brfss_numeric <- brfss_subset %>%
  select(age80, numadult, state, menthlth, checkup1) %>% 
  na.omit()
cor_matrix <- cor(brfss_numeric)
corrplot(cor_matrix, method = "color", type = "upper", tl.cex = 0.8)

```

# Chi Squared iterative test
```{r}
categoricalVars <- names(brfss_subset)[sapply(brfss_subset, is.factor)]
chiResults <- list()

for (i in 1:(length(categoricalVars) - 1)) {
  for (j in (i+1):length(categoricalVars)) {
    var1 <- categoricalVars[i]
    var2 <- categoricalVars[j] # SS - iterate through predictors to compare 2
    
    chiTest <- chisq.test(table(brfss_subset[[var1]], brfss_subset[[var2]]),
                          simulate.p.value = TRUE)
    chiResults[[paste(var1, var2, sep = "_vs_")]] <- # SS - run chi sq and list result
      list(
        variable1 = var1,
        variable2 = var2,
        p_value = chiTest$p.value,
        statistic = chiTest$statistic
      )
  }
}

significantchiResult <- Filter(function(x) x$p_value < 0.05, chiResults)
```
# Chi-squared heatmap and table
```{r}
chiResultsDF <- do.call(rbind, lapply(chiResults, function(x) {
  data.frame(
    variable1 = x$variable1,
    variable2 = x$variable2,
    statistic = x$statistic,
    p_value = x$p_value
  )
}
  )) # SS - do.call allows dynamic execution of another function

significantheatResult <- subset(chiResultsDF, p_value < 0.05)

heatmapData <- acast( #SS - acast to convert to matrix for heatmap
  significantheatResult,
  variable1 ~ variable2,
  value.var =  "statistic",
  fill = NA
)

ggplot(melt(heatmapData, na.rm = TRUE), aes(Var1, Var2, fill = value)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma", name = "Chi-Square Stat") +
  labs(title = "Heatmap of Iterative Chi-Square Stats",
       x = "Variable 1",
       y = "Variable 2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) #SS - create heatmap
```

## Generate Focused table with top results
```{r}
topResultsHeat <- significantheatResult[order(-significantheatResult$statistic), ][1:10, ]
print(topResultsHeat) #SS - Table to show only top 10 results based on statistic and p value less than 0.05
```

#PCA biplot
```{r}
pca_result <- prcomp(brfss_numeric, scale. = TRUE)
biplot(pca_result, main = "PCA Biplot (Centered & Scaled Variables)")
```

# Pearson's Correlation to Measure Numeric
```{r}
cor.test(brfss_subset$age80, brfss_subset$numadult, method = "pearson")
cor.test(brfss_subset$age80, brfss_subset$numadult, method = "spearman")
# Spearman's measurement may provide a more realistic interpretation. Both tests yield statistically significant results.
```


# Generate ggpairs plots to examine collinearity and statistical significance
```{r}
pairsVars <- brfss_subset[, c("age80", "numadult", "state", "menthlth", "checkup1")]
ggpairs(pairsVars)
```

################### descriptive histograms ##################################

```{r}
# Histogram for age80
ggplot(brfss_subset, aes(x = age80)) +
  geom_histogram(binwidth = 5, fill = "darkgreen", color = "white") +
  labs(title = "Age Distribution", x = "Age (Grouped)", y = "Count") +
  theme_minimal()
```

```{r}
# Histogram for numadult
ggplot(brfss_subset, aes(x = numadult)) +
  geom_histogram(binwidth = 1, fill = "purple", color = "white") +
  labs(title = "Number of Adults in Household", x = "Adults", y = "Count") +
  theme_minimal()
```

```{r}
ggplot(brfss_subset, aes(x = ment14d)) +
  geom_density(fill = "skyblue", alpha = 0.6) +
  labs(title = "Density Plot of Poor Mental Health Days", x = "Days", y = "Density") +
  theme_minimal()
```

####################### descriptive statistics ##############################

```{r}
brfss_subset %>%
  select(ment14d, age80, numadult) %>%
  summarise(
    count_ment14d = sum(!is.na(ment14d)),
    mean_ment14d = mean(ment14d, na.rm = TRUE),
    median_ment14d = median(ment14d, na.rm = TRUE),
    sd_ment14d = sd(ment14d, na.rm = TRUE),
    iqr_ment14d = IQR(ment14d, na.rm = TRUE),

    mean_age80 = mean(age80, na.rm = TRUE),
    median_age80 = median(age80, na.rm = TRUE),
    sd_age80 = sd(age80, na.rm = TRUE),
    iqr_age80 = IQR(age80, na.rm = TRUE),

    mean_numadult = mean(numadult, na.rm = TRUE),
    median_numadult = median(numadult, na.rm = TRUE),
    sd_numadult = sd(numadult, na.rm = TRUE),
    iqr_numadult = IQR(numadult, na.rm = TRUE)
  )
```
